Hippocampal segmentation with MAGeT

### 2012-12-05 paper.tex
Fleshed out methods a bit more.. worked in the rationale

\usepackage{alltt}
\usepackage[margin=1.5in]{geometry}
\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable }
\usepackage[square,numbers]{natbib}
\setlength{\marginparwidth}{1.2in}
\newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
\newcommand{\todo}[1]{\marginnote{\textcolor{red}{TODO #1}}}
\newcommand{\mc}[1]{\todo{(MC) #1}}

@@

\subsection{Image Processing and Registration Methods}

Before images were registered, the N3 algorithm [16] is first used to minimize
the intensity nonuniformity in each of the atlases and unlabeled subject
images. In this work we investigated the performance of two registration
methods.

\subsubsection{Automatic Normalization and Image Matching and Anatomical Labeling (ANIMAL)}

The ANIMAL algorithm carries out Image registration is in two phases. In the
first, a 12-parameter linear transformation (3 translations, rotations, scales,
shears) is estimated between images using an algorithm that maximizes the
correlation between blurred MR intensities and gradient magnitude over the
whole brain \citep{Collins}. In the second phase, nonlinear registration is completed using
the ANIMAL algorithm \citep{Collins1995}: an iterative procedure that estimates a 3D
deformation field between two MR images. At first, large deformations are
estimated using blurred version of the input data. These larger deformations
are then input to subsequent steps where the fit is refined by estimating
smaller deformations on data blurred with a Gaussian kernel with a smaller
FWHM. The final transformation is a set of local translations defined on a bed
of equally spaced nodes that were estimated through the optimization of the
correlation coefficient.

For the purposes of this work we used the
regularization parameters optimized in Robbins et al. \citep{Robbins2004}. It
should be noted that the MAGeT brain algorithm is not dependent on this, or
any, particular choice of registration method \citep{Chakravarty2012}.

\todo{should add the MNI website here; also we should add the parameters we used somewhere}

@@

\subsection{Data and Experiments}

Three different experiments were conducted to evaluate the influence of
registration method, atlas set size and resolution, template set size, and
label fusion method and parameters.

\subsubsection{ADNI Validation Experiment}

In this experiment we performed repeated random sub-sampling cross-validation
of the MAGeT algorithm with a pool of 69 images and labels from the ADNI
dataset.  The images are T1-weighted \todo{<insert description of scan/machine
types>}.  The hippocampal labels are provided as part of the ADNI dataset, and
are produced \todo{<insert description of SNT labellings>}.  The parameters we
varied, and the ranges over which we varied are listed in \todo{table}.  We
performed 10 validations per parameter combination.  In each validation trial,
a random sample of subjects were chosen from the pool without replacement as
the atlas set, whereas the template set was chosen randomly with replacement.

\begin{table}
    \begin{tabular}{l|l}
        Parameter           & Values tested                                                      \\ \hline
        Number of Templates & 3 to 20                                                            \\ 
        Number of Atlases   & 3 to 9                                                             \\ 
        Registration Method & ANTS or ANIMAL                                                     \\ 
        Label Fusion Method & majority vote, cross-correlation weighted vote, NMI weighted vote. 
    \end{tabular}
\end{table}

We used the following parameters for ANTS, during registration: 
\begin{verbatim}
  mincANTS 3 -m PR[target_file.mnc,source_file.mnc,1,4] 
   --number-of-affine-iterations 10000x10000x10000x10000x10000 
   --affine-gradient-descent-option 0.5x0.95x1.e-4x1.e-4
   --use-Histogram-Matching --MI-option 32x16000
   -r Gauss[3,0] -t SyN[0.5] -i 100x100x100x20
   -o transformation.xfm
 \end{verbatim}
These settings were adapted from the "reasonable starting point" given in the
ANTS manual. \todo{<rationale?>}

For the purposes of this work we used the regularization parameters optimized
in Robbins et al.[15]. \todo{right? (but added a second layer).}

\todo{multi-atlas comparison, "naive" comparison}

\subsubsection{Winterburn High-resolution Hippocampal Atlas - ADNI Validation}

In this experiment we explored using the MAGeT brain algorithm with the
Winterburn high-resolution atlases to segment 21 randomly chosen MR images from
the ADNI dataset (seven each of healthy, MCI and AD subjects).  The Winterburn
atlases are digital segmentations of the hippocampus in five  in-vivo 300u
isotropic T1-weighted MR scans, and include subfield segmentations for the
cornus ammonis (CA) 1, CA4, dentate gyrus, subiculum, and CA 2 and 3 combined.
Subjects in the Winterburn atlases range in age from 29-57 years (mean age of
37), and include two males and three females.

Since hippocampal segmentation protocols differ between the ADNI labels and
Winterburn atlases, this poses a problem for direct similarity comparisons
between labels produced by MAGeT brain and the ADNI labels.  \todo{explain why
we did(n't) resegment the ADNI images with a the low-res protocol.} To evaluate
the performance of MAGeT brain, we compared classification accuracy of subjects
by diagnosis based on hippocampal volume using both the SMT labels and our
produced labels.  \todo{description of validation -- rms {\tt validate} or
t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV}

% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...) 
%
{\scriptsize\ctable[caption={Descriptive Statistics by Baseline Dx},label=tab,rotate]{lccccc}{\tnote[]{\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.}\tnote[]{Numbers after percents are frequencies.}\tnote[]{\indent Tests used:}\tnote[]{\textsuperscript{\normalfont 1}Kruskal-Wallis test; \textsuperscript{\normalfont 2}Pearson test}}{\FL
\multicolumn{1}{l}{}&\multicolumn{1}{c}{CN}&\multicolumn{1}{c}{LMCI}&\multicolumn{1}{c}{AD}&\multicolumn{1}{c}{Combined}&\multicolumn{1}{c}{Test Statistic}\NN
&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=69$}}&\ML
Age~at~baseline~\hfill\tiny{Years}&{\scriptsize 72.20~}{75.50 }{\scriptsize 78.45} &{\scriptsize 71.00~}{77.10 }{\scriptsize 81.40} &{\scriptsize 71.70~}{77.80 }{\scriptsize 81.75} &{\scriptsize 71.50~}{76.60 }{\scriptsize 81.30} &$ F_{2,66}=0.15 ,~ P=0.862 ^{1} $\NN
Sex~:~Female&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(30)}&$ \chi^{2}_{2}=0 ,~ P=1 ^{2} $\NN
Education&{\scriptsize 16.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 15.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 12.0~}{16.0 }{\scriptsize 16.5} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} &$ F_{2,66}=3.44 ,~ P=0.038 ^{1} $\NN
Marital~status~at~baseline~:~Married&74\%~{\scriptsize~(17)}&87\%~{\scriptsize~(20)}&87\%~{\scriptsize~(20)}&83\%~{\scriptsize~(57)}&$ \chi^{2}_{4}=3.04 ,~ P=0.551 ^{2} $\NN
~~~~Widowed&22\%~{\scriptsize~(~5)}&13\%~{\scriptsize~(~3)}&13\%~{\scriptsize~(~3)}&16\%~{\scriptsize~(11)}&\NN
~~~~Divorced&~4\%~{\scriptsize~(~1)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~1\%~{\scriptsize~(~1)}&\NN
~~~~Never~married&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&\NN
~~~~Unknown&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&\NN
Ethnicity~:~Unknown&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&$^{2}$\NN
~~~~Not~Hisp/Latino&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(69)}&\NN
~~~~Hisp/Latino&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
Race~:~Am~Indian/Alaskan&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&$ \chi^{2}_{2}=5.61 ,~ P=0.061 ^{2} $\NN
~~~~Asian&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
~~~~Hawaiian/Other~PI&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
~~~~Black&~17\%~{\scriptsize~(~4)}&~~4\%~{\scriptsize~(~1)}&~~0\%~{\scriptsize~(~0)}&~~7\%~{\scriptsize~(~5)}&\NN
~~~~White&~83\%~{\scriptsize~(19)}&~96\%~{\scriptsize~(22)}&100\%~{\scriptsize~(23)}&~93\%~{\scriptsize~(64)}&\NN
~~~~More~than~one&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
~~~~Unknown&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
CDR-SB&{\scriptsize 0.00~}{0.00 }{\scriptsize 0.00} &{\scriptsize 0.75~}{1.50 }{\scriptsize 1.50} &{\scriptsize 4.00~}{4.50 }{\scriptsize 5.00} &{\scriptsize 0.00~}{1.50 }{\scriptsize 4.00} &$ F_{2,66}=294.2 ,~ P<0.001 ^{1} $\NN
ADAS~13&{\scriptsize  4.665~}{ 5.670 }{\scriptsize 12.335} &{\scriptsize 14.335~}{16.000 }{\scriptsize 20.500} &{\scriptsize 23.830~}{29.000 }{\scriptsize 31.665} &{\scriptsize 10.000~}{16.000 }{\scriptsize 25.330} &$ F_{2,66}=109.8 ,~ P<0.001 ^{1} $\NN
MMSE&{\scriptsize 28.5~}{29.0 }{\scriptsize 30.0} &{\scriptsize 25.0~}{27.0 }{\scriptsize 28.0} &{\scriptsize 21.0~}{23.0 }{\scriptsize 24.0} &{\scriptsize 24.0~}{27.0 }{\scriptsize 29.0} &$ F_{2,66}=94.17 ,~ P<0.001 ^{1} $\LL
}}

\todo{For minctracc -- we registered to TAL}

\subsection{Results}
\subsubsection{A2A}

Kappa vs. Number of templates: 
Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\maxwidth]{figure/a2a_facet}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\maxwidth]{figure/a2a_facet_mameans}

@@

Kappa minus multi-atlas mean vs. Number of Templates

@@

% latex.default(merge(a_ma_means, t_ma_means, by = "num_atlases"),      colheads = c("Atlases", "ANTS", "ANIMAL"), file = "") 
%
\begin{table}[!tbp]
\begin{center}
\begin{tabular}{lrrr}
\hline\hline
\multicolumn{1}{l}{merge}&\multicolumn{1}{c}{Atlases}&\multicolumn{1}{c}{ANTS}&\multicolumn{1}{c}{ANIMAL}\tabularnewline
\hline
1&$3$&$0.805414417498623$&$0.764266832660852$\tabularnewline
2&$4$&$0.793646605043623$&$0.752420649183594$\tabularnewline
3&$5$&$0.823292141139203$&$0.785546924392891$\tabularnewline
4&$6$&$0.815578947302174$&$0.778345910844141$\tabularnewline
5&$7$&$0.831936342756014$&$0.797761650124922$\tabularnewline
6&$8$&$0.826298462726884$&$0.792443538092187$\tabularnewline
7&$9$&$0.836877632304638$&$0.804568318173359$\tabularnewline
\hline
\end{tabular}
\end{center}
\end{table}

@@

\subsubsection{First Episode Schizophrenic patients}

>>>>

\usepackage{alltt}
\usepackage[margin=1in]{geometry}
\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable }
\usepackage[square,numbers]{natbib}
\setlength{\marginparwidth}{1.2in}
\newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
\newcommand{\todo}[1]{\begin{kframe}{\textcolor{red}{TODO #1}}\end{kframe}}
\newcommand{\mc}[1]{\todo{(MC) #1}}

@@

\todo{Note that algorithm does not specify a registration algorithm or fusion 
method\citep{Chakravarty2012}.  Performance on these is test in experiments}

%%%%%                             Subjects                                 %%%%%

%  Describe subject datasets in general 
\subsection{Subjects}

\subsubsection{ADNI-1 1.5T Screening}
\todo{- image characteristics}
\todo{- describe SNT manual segmentations}
\todo{- baseline/cross-validation dataset demographics?}

% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...) 
%
{\scriptsize\ctable[caption={ADNI-1 1.5T Screening demographics},label=tab,pos=!tbp,]{lrccccc}{\tnote[]{\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.}\tnote[]{$N$\ is the number of non--missing values.}\tnote[]{Numbers after percents are frequencies.}}{\FL
\multicolumn{1}{l}{}&\multicolumn{1}{c}{N}&\multicolumn{1}{c}{CN}&\multicolumn{1}{c}{EMCI}&\multicolumn{1}{c}{LMCI}&\multicolumn{1}{c}{AD}&\multicolumn{1}{c}{Combined}\NN
&&\multicolumn{1}{c}{{\scriptsize $N=414$}}&\multicolumn{1}{c}{{\scriptsize $N=279$}}&\multicolumn{1}{c}{{\scriptsize $N=560$}}&\multicolumn{1}{c}{{\scriptsize $N=278$}}&\multicolumn{1}{c}{{\scriptsize $N=1531$}}\ML
Age~at~baseline~\hfill\tiny{Years}&1531&{\scriptsize 70.9~}{74.0 }{\scriptsize 78.4} &{\scriptsize 66.0~}{71.0 }{\scriptsize 76.2} &{\scriptsize 69.4~}{74.3 }{\scriptsize 79.4} &{\scriptsize 70.9~}{75.9 }{\scriptsize 80.9} &{\scriptsize 69.6~}{74.0 }{\scriptsize 79.2} \NN
Sex~:~Female&1531&50\%~{\scriptsize~(206)}&44\%~{\scriptsize~(123)}&38\%~{\scriptsize~(215)}&45\%~{\scriptsize~(125)}&44\%~{\scriptsize~(669)}\NN
Education&1531&{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 12.0~}{16.0 }{\scriptsize 17.8} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} \NN
Ethnicity~:~Unknown&1531&~1\%~{\scriptsize~(~~~4)}&~1\%~{\scriptsize~(~~~2)}&~1\%~{\scriptsize~(~~~4)}&~1\%~{\scriptsize~(~~~3)}&~1\%~{\scriptsize~(~~13)}\NN
~~~~Not~Hisp/Latino&&96\%~{\scriptsize~(~398)}&95\%~{\scriptsize~(~265)}&96\%~{\scriptsize~(~540)}&96\%~{\scriptsize~(~267)}&96\%~{\scriptsize~(1470)}\NN
~~~~Hisp/Latino&&~3\%~{\scriptsize~(~~12)}&~4\%~{\scriptsize~(~~12)}&~3\%~{\scriptsize~(~~16)}&~3\%~{\scriptsize~(~~~8)}&~3\%~{\scriptsize~(~~48)}\NN
CDR-SB&1531&{\scriptsize 0.0~}{0.0 }{\scriptsize 0.0} &{\scriptsize 0.5~}{1.0 }{\scriptsize 1.5} &{\scriptsize 1.0~}{1.5 }{\scriptsize 2.0} &{\scriptsize 3.5~}{4.5 }{\scriptsize 5.0} &{\scriptsize 0.0~}{1.0 }{\scriptsize 2.5} \NN
ADAS~13&1511&{\scriptsize  6.0~}{ 9.0 }{\scriptsize 12.0} &{\scriptsize  8.0~}{12.0 }{\scriptsize 16.0} &{\scriptsize 14.3~}{18.7 }{\scriptsize 23.0} &{\scriptsize 24.0~}{29.0 }{\scriptsize 34.0} &{\scriptsize 10.0~}{15.3 }{\scriptsize 23.0} \NN
MMSE&1531&{\scriptsize 29.0~}{29.0 }{\scriptsize 30.0} &{\scriptsize 28.0~}{29.0 }{\scriptsize 30.0} &{\scriptsize 26.0~}{27.0 }{\scriptsize 29.0} &{\scriptsize 21.2~}{23.0 }{\scriptsize 25.0} &{\scriptsize 26.0~}{28.0 }{\scriptsize 29.0} \LL
}}

\todo{Check table only includes 1.5T from the Screening}

\subsubsection{SZ First Episode Patients}
\todo{image characteristics, demographics}

% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...) 
%
\begin{table}[!tbp]
\scriptsize
\caption{Schizophrenia First Episode Patient Demographics\label{tab}} 
\begin{center}
\begin{tabular}{lrc}
\hline\hline
\multicolumn{1}{l}{}&\multicolumn{1}{c}{N}&\multicolumn{1}{c}{FEP}\tabularnewline
&&\multicolumn{1}{c}{{\scriptsize $N=81$}}\tabularnewline
\hline
Age&80&{\scriptsize 21~}{23 }{\scriptsize 26} \tabularnewline
Gender~:~M&81&63\%~{\scriptsize~(51)}\tabularnewline
Handedness~:~ambi&81&~6\%~{\scriptsize~(~5)}\tabularnewline
~~~~left&&~5\%~{\scriptsize~(~4)}\tabularnewline
~~~~right&&89\%~{\scriptsize~(72)}\tabularnewline
Education&81&{\scriptsize 11~}{13 }{\scriptsize 15} \tabularnewline
SES~:~lower&81&31\%~{\scriptsize~(25)}\tabularnewline
~~~~middle&&54\%~{\scriptsize~(44)}\tabularnewline
~~~~upper&&15\%~{\scriptsize~(12)}\tabularnewline
FSIQ&79&{\scriptsize  88~}{102 }{\scriptsize 109} \tabularnewline
\hline
\end{tabular}
\end{center}
\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.\\$N$\ is the number of non--missing values.\\Numbers after percents are frequencies.\end{table}

%%%%%                       Image pre-processing                           %%%%%

\subsection{Image pre-processing}

Before images were registered, the N3 algorithm [16] is first used to minimize
the intensity nonuniformity in each of the atlases and unlabeled subject
images.

%%%%%                       Registration Methods                           %%%%%

\subsection{Registration}

\subsubsection{Automatic Normalization and Image Matching and Anatomical Labeling (ANIMAL)}
The ANIMAL algorithm carries out Image registration is in two phases. In the
first, a 12-parameter linear transformation (3 translations, rotations, scales,
shears) is estimated between images using an algorithm that maximizes the
correlation between blurred MR intensities and gradient magnitude over the
whole brain \citep{Collins}. In the second phase, nonlinear registration is completed using
the ANIMAL algorithm \citep{Collins1995}: an iterative procedure that estimates a 3D
deformation field between two MR images. At first, large deformations are
estimated using blurred version of the input data. These larger deformations
are then input to subsequent steps where the fit is refined by estimating
smaller deformations on data blurred with a Gaussian kernel with a smaller
FWHM. The final transformation is a set of local translations defined on a bed
of equally spaced nodes that were estimated through the optimization of the
correlation coefficient.

For the purposes of this work we used the regularization parameters optimized 
in Robbins et al. \citep{Robbins2004}.

\todo{link to MNI website here}
\todo{command line}

@@

%%%%%                            Experiments                               %%%%%

\subsection{Experiments}

Experiments were performed to assess the performance of MAGeT brain with various 
parameter settings as well as on diverse datasets.  In each experiment we contrast 
the performance of MAGeT brain with that standard single- and multi-atlas 
segmentations derived from the same atlas library.

%%%%%
%%%%%                 Experiments: ADNI-1 cross-validationn               %%%%%
%%%%%
\subsubsection{ADNI-1 cross-validation}

To test the accuracy of the MAGeT brain algorithm with different parameter 
settings, repeated random sub-sampling cross-validation (RRSCV) was performed
on a subset of the ADNI-1 dataset.

{\bf Dataset evaluated.} 69 1.5T images were randomly selected from the 
{\em ADNI1:Screening 1.5T} standardized dataset.  Demographics for this subset 
are shown in Table X\todo{FIXME: table reference}.

% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...) 
%
{\scriptsize\ctable[caption={ADNI-1 cross-validation subset demographics},label=tab,pos=!tbp,]{lcccc}{\tnote[]{\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.}\tnote[]{Numbers after percents are frequencies.}}{\FL
\multicolumn{1}{l}{}&\multicolumn{1}{c}{CN}&\multicolumn{1}{c}{LMCI}&\multicolumn{1}{c}{AD}&\multicolumn{1}{c}{Combined}\NN
&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=69$}}\ML
Age~at~baseline~\hfill\tiny{Years}&{\scriptsize 72.2~}{75.5 }{\scriptsize 78.5} &{\scriptsize 71.0~}{77.1 }{\scriptsize 81.4} &{\scriptsize 71.7~}{77.8 }{\scriptsize 81.8} &{\scriptsize 71.5~}{76.6 }{\scriptsize 81.3} \NN
Sex~:~Female&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(30)}\NN
Education&{\scriptsize 16.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 15.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 12.0~}{16.0 }{\scriptsize 16.5} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} \NN
Ethnicity~:~Unknown&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}\NN
~~~~Not~Hisp/Latino&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(69)}\NN
~~~~Hisp/Latino&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}\NN
CDR-SB&{\scriptsize 0.00~}{0.00 }{\scriptsize 0.00} &{\scriptsize 0.75~}{1.50 }{\scriptsize 1.50} &{\scriptsize 4.00~}{4.50 }{\scriptsize 5.00} &{\scriptsize 0.00~}{1.50 }{\scriptsize 4.00} \NN
ADAS~13&{\scriptsize  4.67~}{ 5.67 }{\scriptsize 12.34} &{\scriptsize 14.34~}{16.00 }{\scriptsize 20.50} &{\scriptsize 23.83~}{29.00 }{\scriptsize 31.66} &{\scriptsize 10.00~}{16.00 }{\scriptsize 25.33} \NN
MMSE&{\scriptsize 28.5~}{29.0 }{\scriptsize 30.0} &{\scriptsize 25.0~}{27.0 }{\scriptsize 28.0} &{\scriptsize 21.0~}{23.0 }{\scriptsize 24.0} &{\scriptsize 24.0~}{27.0 }{\scriptsize 29.0} \LL
}}

{\bf Atlas and template libary.}  Atlases consisted of images taken from the 
dataset, with corresponding manual labels provided by SNT.  Atlas library size
was varied from 3 to 9 images.  The remaining images were segmented, with the 
template library size varying from 3 to 20 images.  Template library images were
selected randomly from the images to be segmented.

{\bf Registration method.}  Both the ANTS and ANIMAL registration methods were 
used.
\todo{In ANIMAL, images aligned to TAL space}

{\bf Label fusion.} Majority vote, Cross-correlation weighted majority vote, and
Normalized Mutual Information weighted majority vote are used.  With the 
weighted majority vote fusion methods, the number of top labels used in the 
fusion was varied from 3 to 20 images.

{\bf Evaluation.}  Repeated random sub-sampling cross-validation (RRSCV) 
consists of repeated trials in which items from the dataset are randomly assigned to a training set or validation set. In each trial, performance on the validation set is measured, and then averaged across all trials.

We performed RRSCV on each combination of parameters listed above: atlas library size, template library size, registration method, and label fusion method.  We performed 10 trials per parameter combination.  In each validation trial, the training set consisted of the images used as atlases, and the validation set consisted of the images to be segmented. The MAGeT brain algorithm and the basic multi-atlas segmentation procedure were applied to segment the images in the validation set.  Additionally, in each trial, the single-atlas segmentation was obtained for each atlas-template.

The gold-standard for the segmentation accuracy of images in the validation set was the SNT manual labels.

\todo{ number of registrations/comparisions }

%%%%%
%%%%%               Experiments: ADNI-1 Screening Validation               %%%%%
%%%%%
\subsubsection{ADNI-1 Screening Validation}

To test the accuracy of MAGeT brain on a real-world task we segment the entire
ADNI-1 dataset using an atlas set that is not representative of the subject set.

{\bf Dataset evaluated.} All images from the {\em ADNI1:Screening 1.5T} 
standardized dataset.

{\bf Atlas and template libary.} The atlas library consisted of the entire 
Winterburn atlas set.  The Winterburn atlases are digital segmentations of the
hippocampus in five in-vivo 300u isotropic T1-weighted MR scans, and include 
subfield segmentations for the cornus ammonis (CA) 1, CA4, dentate gyrus, 
subiculum, and CA 2 and 3 combined. Subjects in the Winterburn atlases range in 
age from 29-57 years (mean age of 37), and include two males and three females.

The template library consisted of 21 randomly selected images from the ADNI1 
data dataset (7 healthy, MCI and AD subjects).

{\bf Registration method.}  - ANTS as it performed best in the cross-validation
experiment. \todo{how to reference results whilst still in Methods?}

{\bf Label fusion.} Majority vote, as it is simpliest to run and performed 
equally well in cross-validation experiment.

{\bf Evaluation.}

Since hippocampal segmentation protocols differ between the ADNI labels and
Winterburn atlases, this poses a problem for direct similarity comparisons
between labels produced by MAGeT brain and the ADNI labels.  \todo{explain why
we did(n't) resegment the ADNI images with a the low-res protocol.}

To evaluate the performance of MAGeT brain, we correlate our segmentation
volumes with manual segmentation volumes, as well as with hippocampal volumes
of established automated segmentation methods.

Additionally, we compared classification accuracy of subjects
by diagnosis based on hippocampal volume using both the SMT labels and our
produced labels.  \todo{description of validation -- rms {\tt validate} or
t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV}

%%%%%
%%%%%          Experiments: SZ First Episode Patient Validation           %%%%%
%%%%%
\subsubsection{SZ First Episode Patient Validation}
 To validate that MAGeT's performance generalises to other diseases.
Measure performance using best parameter settings on a different disease dataset.

{\bf Dataset evaluated.} SZ First Episode Patients

{\bf Atlas and template libary.} two different atlas sets: a manual hippocampal
segmentation of patients, and Winterburn atlas set.

{\bf Registration method.} ANTS.

{\bf Label fusion.} Majority vote.

{\bf Evaluation.} We validate the FEP-atlas segmentations using Dice's Kappa,
and the Winterburn-atlas segmentations by correlating volumes.

%%%%%
%%%%%              Experiments: Winterburn Atlases Validation              %%%%%
%%%%%
\subsubsection{Winterburn Atlases Validation}
{\bf Dataset evaluated.} T1 BRAVO scans of the same subjects included in the 
Winterburn atlas set.  These scans are taken within .... weeks of the scans for 
the Winterburn atlases.

{\bf Atlas and template libary.} Atlas library is Winterburn T1 atlases.  
Template library consists of all five T1 BRAVOs, plus 15 T1 healthy control images.

{\bf Registration method.} ANTS

{\bf Label fusion.} Majority vote.

{\bf Evaluation.}  Leave one out cross-validation (LOOCV) in which all five
subjects are segmented in separate runs of MAGeT brain.  In each run, the subject
to be segmented is excluded from the Atlas library (so only four atlases are used).

Segmentation accuracy is judged by difference in hippocampal volume.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                               Results
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Results}

%%%%%
%%%%%                  Results: ADNI-1 cross-validationn                   %%%%%
%%%%%
\subsubsection{ADNI-1 Cross-Validation}

 - find significant improvement over multi-atlas performed with the same
  parameters. Also, find smoothed performance is monotonically increasing but asymptotic
  in size of both template and atlas library, with peak performance reached after 15
  templates. \todo{Can we statistically capture "peak" performance?  The point at which 
  gains become statistically insignificant?}

\begin{figure}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\maxwidth]{figure/ADNI1_cross_validation_means}

@@

  \caption{Comparison of MAGeT performance on ADNI-1 subset.  Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))}
  \label{ADNI1_cross_validation_means}
\end{figure}

\begin{figure}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\maxwidth]{figure/ADN1-cross-validation-MA-difference}

@@

  \caption{Difference in mean Kappa between MAGeT brain and multi-atlas}
  \label{ADNI1_cross_validation_MA_difference}
\end{figure}

@@

% latex.default(merge(a_ma_means, t_ma_means, by = "num_atlases"),      colheads = c("Atlases", "ANTS", "ANIMAL"), file = "", dec = 2) 
%
\begin{table}[!tbp]
\begin{center}
\begin{tabular}{lrrr}
\hline\hline
\multicolumn{1}{l}{merge}&\multicolumn{1}{c}{Atlases}&\multicolumn{1}{c}{ANTS}&\multicolumn{1}{c}{ANIMAL}\tabularnewline
\hline
1&$3.00$&$0.81$&$0.76$\tabularnewline
2&$4.00$&$0.79$&$0.75$\tabularnewline
3&$5.00$&$0.82$&$0.79$\tabularnewline
4&$6.00$&$0.82$&$0.78$\tabularnewline
5&$7.00$&$0.83$&$0.80$\tabularnewline
6&$8.00$&$0.83$&$0.79$\tabularnewline
7&$9.00$&$0.84$&$0.80$\tabularnewline
\hline
\end{tabular}
\end{center}
\end{table}

@@

\subsubsection{First Episode Schizophrenic patients}
High volume correlation between Winterburn segmentation volumes and ground truth.  (High-ish?) Kappa when using manual segmentations as Atlases.

\begin{figure}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
